Place of death in NRS COVID-19 deaths: data processing for analysis, 2022 data

### 2022-03-07 covid_nrs_place_of_death_data_2022.Rmd
Modified code to account for missing "crude_positive_rate" in new case trends data - the rate is now computed from population estimate and cumulative cases; also included the "functions.R" file - forgot to commit earlier!

```{r}
weekly_covid_cases_hb <-
  cases_hb %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive),
    cumulative_positive = max(cumulative_positive),
    crude_rate_positive_per100k = max(crude_rate_positive),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )

weekly_covid_cases_la <-
  cases_la %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive),
    cumulative_positive = max(cumulative_positive),
    crude_rate_positive_per100k = max(crude_rate_positive),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )
```

@@

>>>>

@@

## Calculating population size of areas

Update 7 Mr 2022: at some point the file structure changed, and the crude positive rate per 100k was removed... however, the crude mortality rate remains, so we can infer the population from the `crude_rate_deaths` and `cumulative_deaths` figures, and compute the `crude_rate_positive` from those.

The trouble is that early on in the data, when `cumulative_deaths` and `crude_rate_deaths` are both 0, we can't infer the population size. If the population size is constant in the data, we can calculate it for each `ref_area` (HB or LA) and use that as a lookup table.

It looks like the same population estimate was used for data 2020-2022, and it is the **mid-year 2019 estimate**.

## Importing & calculating case trends

```{r}
weekly_covid_cases_hb <-
  cases_hb %>%
  left_join(
    population_estimate_hb_2019 %>%
    add_scotland_total_to_population_estimate %>%
    select(ref_area = hb, population_size = all_ages)
    ) %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive, na.rm = TRUE),
    cumulative_positive = max(cumulative_positive, na.rm = TRUE),
    # crude_rate_positive_per100k = max(crude_rate_positive),  # this column is no longer in the data, so we compute it below...
    crude_rate_positive_per100k = max(cumulative_positive / population_size * 1e5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )

weekly_covid_cases_la <-
  cases_la %>%
  left_join(
    population_estimate_la_2019 %>%
    add_scotland_total_to_population_estimate %>%
    select(ref_area = la, population_size = all_ages)
    ) %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive, na.rm = TRUE),
    cumulative_positive = max(cumulative_positive, na.rm = TRUE),
    # crude_rate_positive_per100k = max(crude_rate_positive),  # this column is no longer in the data, so we compute it below...
    crude_rate_positive_per100k = max(cumulative_positive / population_size * 1e5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )
```
